Debugging by \textit{lastChange}

### 2010-11-18 papers/pldi2011/sigplanconf-template.tex

\documentclass[preprint]{sigplanconf}

@@

\authorinfo{Salman Mirghasemi}
           {Ecole Polytechnique F\'ed\'erale de Lausanne}
           {salman.mirghasemi@epfl.ch}

\authorinfo{John J. Barton}
           {IBM Research - Almaden}
           {johnjbarton@johnjbarton.com}

\authorinfo{Claude Petitpierre}
           {Ecole Polytechnique F\'ed\'erale de Lausanne}
           {claude.petitpierre@epfl.ch}

@@

Developers often seek the origins of wrong values they see in their
debugger. Their search must be backwards in time: the code causing the
wrong value executed before the wrong value appeared. Searching with
breakpoint- or log- based debuggers demands persistence and
significant experience with the application being debugged. We propose
a new approach, an algorithm called \textit{lastChange}, which automatically
locates the last point that a value was changed and related
information such as the call stack and program state. Starting from a
program halted on a breakpoint, the \textit{lastChange} solution
applies queries to the live program during re-execution. Thus
\textit{lastChange} combines the flexibility of breakpoint debugging
with the expressive power of log-based query debugging.  Contrary to
other replay-based algorithms which require exactly the same
re-executions, \textit{lastChange} only requires \textit{bug
  reproducibility}, a test case is available which reproduces the bug.

As a proof of concept, we developed \textit{Querypoint}, a prototype
which enhances a popular JavaScript debugger, Firebug, with
\textit{lastChange} feature. Moreover, \textit{Querypoint} provides
mechanisms for automated bug reproduction, and a novel user interface
to investigated execution points and results.

@@

According to \cite{LaToza}, developers spend about fifty percent of
their time debugging. To fix a bug, developers typically reproduce and
monitor the buggy execution several times to understand the program's
unexpected behavior. Trial and error, guess-work, and analyzing large
and complicated data make debugging difficult and time-consuming. Any
enhancement to debugging techniques and tools can save developers'
time, reduce development costs and improve programs' quality.

@@

To address these problems, we introduce a new functionality in
debuggers, \textit{lastChange}, which locates the origin of a wrong
value. It uses program re-execution similar to the practice
in using breakpoints and log statements.

Imagine a program execution paused on a breakpoint and a developer
suspicious about the value of an object property or a variable. The
developer queries \textit{lastChange} on the value. The debugger
reproduces the buggy execution and collects data
and once the execution reaches the same place (i.e., the same
breakpoint hit), it pauses the execution, analyzes the collected data
and shows the location of the last change to the developer. The
developer can also examine the program state at the located point of
execution, and continue debugging by more \textit{lastChange} queries
from that point.

@@

Our contribution in this paper is the algorithm \textit{lastChange},
which locates the last place a value has changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The algorithm builds on existing breakpoint debugger
technology. We demonstrate the feasibility of the algorithm with
\textit{Querypoint}, an implementation extending the Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.

The rest of the paper is organized as follows. First, we demonstrate
the \textit{lastChange} usage on a simple example with the comparison
to breakpoint debugging. Section 3 presents \textit{lastChange}
algorithm. In section 4, we explain the details of the JavaScript
prototype implementation. We discuss the effect of non-determinism on
the \textit{lastChange} algorithm. Section 6 contains an evaluation of
\text{lastChange} functionality.

@@

>>>>

\documentclass{sigplanconf} %[preprint]

@@

%\authorinfo{Salman Mirghasemi}
%           {Ecole Polytechnique F\'ed\'erale de Lausanne}
%           {salman.mirghasemi@epfl.ch}

%\authorinfo{John J. Barton}
%           {IBM Research - Almaden}
%           {johnjbarton@johnjbarton.com}

%\authorinfo{Claude Petitpierre}
%           {Ecole Polytechnique F\'ed\'erale de Lausanne}
%           {claude.petitpierre@epfl.ch}

\authorinfo{First Author}
           {AAA university}
           {a@a.com}

\authorinfo{Second Author}
           {BBB company}
           {b@b.com}

\authorinfo{Third Author}
           {CCC university}
           {c@c.com}

@@

Developers often seek the origins of wrong values they see in their
debugger. Their search must be backwards in time: the code causing the
wrong value executed before the wrong value appeared. Searching with
breakpoint- or log- based debuggers demands persistence and
significant experience with the application being debugged. We propose a new 
practical technique called \textit{lastChange}, which automatically
locates the last point that a variable or an object property  was changed and provides related
information such as the call stack and program state. Starting from a
program halted on a breakpoint, the \textit{lastChange} solution
applies queries to the live program during re-execution. Thus
\textit{lastChange} combines the flexibility of breakpoint debugging
with the expressive power of log-based query debugging.  Contrary to
other replay-based approaches which require exactly the same
re-executions (fully deterministic), \textit{lastChange} only requires \textit{bug 
reproducibility}, a test case is available which reproduces the bug.

%TODO these paragraph should contain the summary of evaluation results.
As a proof of concept, we developed \textit{Querypoint}, a prototype
which enhances a popular JavaScript debugger, Firebug, with
\textit{lastChange} feature. Moreover, \textit{Querypoint} provides
mechanisms for automated bug reproduction, and a novel user interface
to investigated execution points and results.

@@

According to \cite{LaToza}, developers spend about fifty percent of
their time debugging. Trial and error, guess-work, and analyzing large
and complicated data make debugging difficult and time-consuming. To fix 
a bug, developers typically reproduce and monitor the buggy execution several 
times to understand the program's unexpected behavior. Therefore, any
enhancement to debugging techniques and tools which automates or facilitates developers' actions
on execution traces, can save developers' time, reduce development costs and improve programs' quality.

@@

To address these problems, we introduce a new functionality in
debuggers, \textit{lastChange}, which locates the origin of a wrong
value. It uses program re-execution similar to the practice
in using breakpoints and log statements. Imagine a program execution 
paused on a breakpoint and the developer is
suspicious about the value of a variable or an object property. The
developer queries \textit{lastChange} on the value. The debugger
reproduces the buggy execution and collects data
and once the execution reaches the same place (i.e., the same
breakpoint hit), it pauses the execution, analyzes the collected data
and shows the location of the last change to the developer. The
developer can also examine the program state at the located point of
execution, and continue debugging by more \textit{lastChange} queries
from that point.

@@

Our contribution in this paper is the technique \textit{lastChange},
which locates the last place a value has changed, gathers other values
from that execution point, and allows \textit{lastChange} operations
from that point. The technique builds on existing breakpoint debugger
technology. We demonstrate the feasibility of the approach with
\textit{Querypoint}, an implementation extending the Firebug
JavaScript debugger. \textit{Querypoint} also provides mechanism for
automated bug reproduction, and a novel user interface which
summarizes investigated execution points and collected results.

The rest of the paper is organized as follows. First, we demonstrate
the \textit{lastChange} usage on a simple example with the comparison
to breakpoint debugging. Section 3 presents \textit{lastChange}
algorithm. In section 4, we explain the details of the JavaScript
prototype implementation. We discuss the effect of non-determinism on
the \textit{lastChange} results in section 5. %TODO update it according to new
Section 6 contains an evaluation of\text{lastChange} functionality.

@@

\section{Evaluation}
%To verify that lastChange can be effective we performed a small study of developers debugging with and without lastChange. Our goal is small, only to show that one aspect of debugging, the determination the causes of incorrect values, can be improved. The broader impact of lastChange on the total debugging process will have to wait for implementation of the functions described in section XX directly in a Web browser JavaScript engine.

%Four experienced Javascript developers used the Firebug debugger on four small programs we provided. Each developer first used breakpoint debugging on two programs. Then we had them read sections YY - ZZ of this paper. Finally each developer applied the Querypoint prototype to two more programs. The programs were preselected so each developer saw them in a different order. Figure AA compares the human and computing times for these tests. [results]

% Of course these programs were designed to highlight lastChange and all kinds of debugging issues have been hidden by the design of our tests. Our results show that when a defect relates to incorrect values and a developer recognizes that this, then the operational mechanics of lastChange lead to the defect much more quickly than breakpoints.
